Real-time Attention Assessment for Social Robots

### 2015-09-15 paper.tex
Working on the intro/state of art

\section{Introduction: Attention and Engagement}
%In the recent years, research is pushing forward in building intelligent systems
%and environments that try to maximize the benefits of cooperation with humans
%based on verbal and non-verbal communication. The interaction between social
%robots and humans entails a broad set of situations, and evidences the need for
%capturing significant backchannel information generated by the user during the
%interaction process. Such information is used to adapt the system to the
%specific user's needs to eventually provide a social or emotional intelligence
%to the agents.
%
%It is indeed not enough to exhibit human-like capabilities in social tasks: a
%robot needs to achieve human-like manners. For instance, several studies showed
%that humans prefer to interact with robots whose task outcome is delayed, shows
%uncertainty or an undesired output \cite{Admoni,Short}. In fact, the ability to
%adapt has been proved to be beneficial in long-term interactions
%\cite{Tielman:2014, Lim:2014}, specifically with children, manifesting a more
%positive attitude over time.

In order to build intelligent agents, it is necessary to equip them with a range
of perceptual capabilities: while face and object tracking and
recognition~\cite{Zhao:2003, Jafri:2014}, path planning~\cite{Galceran:2013},
speech recognition~\cite{brick2007incremental} or task
learning~\cite{calinon2007learning} are amongst the current major research
directions, existing literature tends to evaluate each of these algorithms in
their own metric space, without considering the overall interaction quality.
Along this line, Anzalone \etal~\cite{anzalone} have recently argued in favor of
such a global evaluation, and propose to assess these algorithms in term of
their capability to obtain the desired effect in a human-robot interaction
context, and they correspondingly propose metrics build around the measurement
of \emph{engagement} as indicator of the quality of the experience.

Engagement describes a cognitive, affective (specifically intrinsic motivation),
and behavioral state of interaction with a computer application that ``makes the
user want to be there" \cite{OBrien:2010}. Engaging interactions were thought to
involve intrinsic interest, interactivity, perceived control and choice,
functionality, motivation and attention. In fact,
\cite{yarbus1967eye,barber1976perception} suggest that human's attention is
highly correlated with what they look at, and the existing relationship between
gaze and attention during social interaction has been exposed
in~\cite{argyle1969social}. In this way, in situ Visual Focus of Attention
(VFoA) estimation is advantageous to human-robot interaction and a major
contribution towards a better understanding of the interaction quality.

@@

\subsection*{Engagement and Attention: Related Work}

@@

Technique for head pose estimation: based on depth
sensing~\cite{fanelli2012real}.

In our work, we rely on recent advances on template-based face
alignement~\cite{kazemi2014one} that allows fast (in the order of
a few milliseconds) face features extraction on 2D images, that we found to be fairly
robust to real-world lighting conditions and occlusions naturally occuring in child-robot
interactions.

>>>>

\section{Introduction}

%In the recent years, research is pushing forward in building intelligent systems
%and environments that try to maximize the benefits of cooperation with humans
%based on verbal and non-verbal communication. The interaction between social
%robots and humans entails a broad set of situations, and evidences the need for
%capturing significant backchannel information generated by the user during the
%interaction process. Such information is used to adapt the system to the
%specific user's needs to eventually provide a social or emotional intelligence
%to the agents.
%
%It is indeed not enough to exhibit human-like capabilities in social tasks: a
%robot needs to achieve human-like manners. For instance, several studies showed
%that humans prefer to interact with robots whose task outcome is delayed, shows
%uncertainty or an undesired output \cite{Admoni,Short}. In fact, the ability to
%adapt has been proved to be beneficial in long-term interactions
%\cite{Tielman:2014, Lim:2014}, specifically with children, manifesting a more
%positive attitude over time.

In order to build intelligent agents, it is necessary to equip them with a range
of perceptual capabilities: while face and object tracking and
recognition~\cite{Zhao:2003, Jafri:2014}, path planning~\cite{Galceran:2013},
speech recognition~\cite{brick2007incremental} or task
learning~\cite{calinon2007learning} are amongst the current major research
directions, existing literature tends to evaluate each of these algorithms in
their own metric space, without considering the overall interaction quality.
Along this line, Anzalone~\etal~\cite{anzalone} have recently argued in favor of
such a global evaluation, and propose to assess these algorithms in term of
their capability to obtain the desired effect in a human-robot interaction
context, and they correspondingly propose metrics build around the measurement
of \emph{engagement} as indicator of the quality of the experience.

Engagement describes a cognitive, affective (specifically intrinsic motivation),
and behavioral state of interaction with a computer application that ``makes the
user want to be there" \cite{OBrien:2010}. Engaging interactions were thought to
involve intrinsic interest, interactivity, perceived control and choice,
functionality, motivation and attention.

@@

\subsection*{Attention and Engagement: Related Work}

@@

The relations between one's focus of attention and what he/she is looking at has
long been established~\cite{yarbus1967eye,barber1976perception}, and more
specifically, the existing relationship between gaze and attention during social
interaction and the related gaze patterns, belongs to classic textbooks
like~\cite{argyle1969social}. As such, measuring the direction of gaze is
a natural proxy to estimate the (visual) focus of attention of a social agent.

However, performing such a measure on a robot, in real-time, and in ecologically
valid environments (which rules out bulky or invasive apparatus like
eye-trackers, or techniques requiring fine calibration and/or static
interactions) remains a key challenge in the field of Human-Robot Interaction.

The current state-of-the-art techniques found in HRI that operate both on-line
and during field experiments rely on head pose
estimation, most often based on depth sensors (RGB-D). Fanelli~\etal provides a
overview of these approaches in~\cite{fanelli2012real}. While fairly robust, these
techniques remain CPU intensive for a limited accuracy.

In our work, we rely on recent advances on template-based face
alignment~\cite{kazemi2014one} that allows fast (in the order of a few
milliseconds) face features extraction on 2D images, that we found to be fairly
robust to real-world lighting conditions and occlusions naturally occurring in
child-robot interactions.

In this way, in situ Visual Focus of Attention
(VFoA) estimation is advantageous to human-robot interaction and a major
contribution towards a better understanding of the interaction quality.
